

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Preprocessing of data (replace NA values by mean and delete columns having NA mean and not used for building models)

```{r preprocessing}
LoanData <- read.csv("C:/Users/DELL/Downloads/UMCP/Pi Analytics/loan_data.csv")

for(i in 1:ncol(LoanData)){
  LoanData[is.na(LoanData[,i]), i] <- mean(LoanData[,i], na.rm = TRUE)
}

na_columns <- colnames(LoanData)[ apply(LoanData, 2, anyNA) ]

unique(LoanData$title)
mean(LoanData$title)

LoanData$emp_title <- NULL
LoanData$title <- NULL
LoanData$verification_status_joint <- NULL

```

#Create a new column with mid of fico band (this is to be used for building models)

```{r create a new column with mid of fico band to be used for building models}
library(stringr)
appl_fico_low <- str_split_fixed(LoanData$APPL_FICO_BAND, "-", 2)
#appl_fico_low
last_fico_low <- str_split_fixed(LoanData$Last_FICO_BAND, "-", 2)
#last_fico_low
last_fico_low[last_fico_low == "LOW"] <- 499
last_fico_low[last_fico_low == "MISSING"] <- 499
last_fico_low[last_fico_low == ""] <- 499
appl_fico_low_lower <- as.numeric(appl_fico_low[,1])
appl_fico_low_upper <- as.numeric(appl_fico_low[,2])
last_fico_low_lower <- as.numeric(last_fico_low[,1])
last_fico_low_upper <- as.numeric(last_fico_low[,2])
last_fico_low_upper[is.na(last_fico_low_upper)] <- 499
#str(last_fico_low_upper)
change_fico_lower <- appl_fico_low_lower - last_fico_low_lower
change_fico_upper <- appl_fico_low_upper - last_fico_low_upper
#summary(change_fico_lower)
FICO_change <- paste(change_fico_lower, change_fico_upper, sep="-")
#FICO_change

appl_fico <- data.frame(appl_fico_low_lower,appl_fico_low_upper)
mean_appl_fico <- rowMeans(appl_fico)
last_fico <- data.frame(last_fico_low_lower,last_fico_low_upper)
mean_last_fico <- rowMeans(last_fico)
change_fico <- data.frame(change_fico_lower,change_fico_upper)
mean_change_fico <- rowMeans(change_fico)

LoanData <- data.frame(LoanData,mean_appl_fico,mean_change_fico,mean_last_fico)
```

#Sampling the data

```{r sampling}
set.seed(123457)

train<-sample(nrow(LoanData),0.7*nrow(LoanData))
data_train<-LoanData[train,]
data_val<-LoanData[-train,]
```

#Multiple linear regression

```{r linear regression}
fit <- lm(InterestRate~MonthlyIncome+HomeOwnership+RevolvingLineUtilization+EmploymentLength+mean_appl_fico+funded_amnt+verification_status,data=data_train)
summary(fit)
plot(fit)

actual <- data_val$InterestRate
predict <- predict(fit, newdata=data_val)

Metrics <- c("AE","RMSE","MAE","SSE","MSE")
x1 <- mean(actual - predict)
x2 <- sqrt(mean((actual - predict)^2))
x3 <- mean(abs(actual - predict))
x4 <- sum(actual - predict)^2
x5 <- mean((actual - predict)^2)
Values <- c(x1,x2,x3,x4,x5)
x <- data.frame(Metrics,Values)
x
```

##Remove factors greater than 32 levels since trees dont run for these variables

```{r remove factors that have more than 32 levels for building regression tree}
str(LoanData)
is.fact <- sapply(LoanData, is.factor)
factors.df <- LoanData[, is.fact]
lapply(factors.df, levels)

LoanData$RECEIVED_D <- NULL
LoanData$MONTH <- NULL
LoanData$IssuedDate <- NULL
LoanData$State <- NULL
LoanData$EarliestCREDITLine <- NULL
LoanData$APPL_FICO_BAND <- NULL
LoanData$Last_FICO_BAND <- NULL
LoanData$VINTAGE <- NULL
LoanData$int_rate <- NULL
LoanData$sub_grade <- NULL
LoanData$issue_d <- NULL
LoanData$url <- NULL
LoanData$zip_code <- NULL
LoanData$desc <- NULL
LoanData$addr_state <- NULL
LoanData$earliest_cr_line <- NULL
LoanData$last_pymnt_d <- NULL
LoanData$next_pymnt_d <- NULL
LoanData$last_credit_pull_d <- NULL
LoanData$emp_length <- NULL
LoanData$loan_status <- NULL
```

#OR (since not running for full set of variables) create new dataframe using only selected variables

```{r new dataframe for regression tree}
LoanDataNew <- data.frame(LoanData$InterestRate,LoanData$MonthlyIncome,LoanData$HomeOwnership,LoanData$RevolvingLineUtilization,LoanData$EmploymentLength,LoanData$mean_appl_fico,LoanData$funded_amnt,LoanData$verification_status)
```

#Resample using new dataframe

```{r sample from new dataframe}
set.seed(123457)

train<-sample(nrow(LoanDataNew),0.7*nrow(LoanDataNew))
data_train<-LoanDataNew[train,]
data_val<-LoanDataNew[-train,]
```

#Regression tree

```{r regression tree}
library(MASS)
library(tree)
tree.LoanData.regression = tree (LoanData.InterestRate~.,data_train)
summary(tree.LoanData.regression)
tree.LoanData.regression
plot(tree.LoanData.regression)
text(tree.LoanData.regression,pretty=0)
```

#Pruned regression tree

```{r pruned regression tree}
prune.LoanData.regression = prune.tree(tree.LoanData.regression,best=6)
plot(prune.LoanData.regression)
text(prune.LoanData.regression,pretty=0)
```

#Prediction using regression tree and MSE

```{r Prediction using regression tree and MSE}
pred.LoanData.prune.regression = predict(prune.LoanData.regression,data_val)
plot(pred.LoanData.prune.regression,data_val$LoanData.InterestRate)
abline(0,1)
mean((pred.LoanData.prune.regression-data_val$LoanData.InterestRate)^2)

table(pred.LoanData.prune.regression)
```

#Bagging
```{r bagging}
library(randomForest)
set.seed(123457)
#Bagging is just random forest with m = no. of predictors
bag.LoanData=randomForest(LoanData.InterestRate~.,data=data_train,mtry=7,importance=TRUE)
bag.LoanData
yhat.bag = predict(bag.LoanData,newdata=data_val)
yhat.bag
plot(yhat.bag, data_val$InterestRate)
abline(0,1)
mean((yhat.bag-data_val$InterestRate)^2)
yhat.test=data_val$InterestRate
yhat.test
importance(bag.LoanData)
varImpPlot(bag.LoanData)
```

#Random forest
```{r random forest}
set.seed(123457)
rf.LoanData=randomForest(LoanData.InterestRate~.,data=data_train,mtry=2,importance=TRUE)
rf.LoanData
yhat.rf = predict(rf.LoanData,newdata=data_val)
yhat.rf
mean((yhat.rf-data_val$InterestRate)^2)
yhat.test = data_val$InterestRate
yhat.test
importance(rf.LoanData)
varImpPlot(rf.LoanData)
```

#Boosting
```{r boosting}
#install.packages("gbm")
library(gbm)
set.seed(123457)
boost.CreditData=gbm(InterestRate~.,data=data_train,distribution="gaussian",n.trees=5000,interaction.depth=10)
summary(boost.LoanData)
yhat.boost=predict(boost.LoanData,newdata=data_val,n.trees=5000)
yhat.boost
mean((yhat.boost-data_val$InterestRate)^2)
yhat.test=data_val$InterestRate
yhat.test
```

#Kmeans (exploratory analysis)
```{r kmeans}
set.seed(123457)
library(dummies)
str(LoanDataNew)
LoanDatakmeans <- dummy.data.frame(LoanDataNew, names = c("LoanData.HomeOwnership","LoanData.EmploymentLength","LoanData.verification_status") ,sep = ".")
LoanDataScale <- scale(LoanDatakmeans)
km.out <- kmeans(LoanDataScale,8,nstart=20)
km.out
km.out$centers
dist(km.out$centers)
cluster <- km.out$cluster
cluster

LoanDataCluster <- data.frame(LoanDataNew,km.out$cluster)
LoanDataCluster <- data.frame(LoanDataCluster,LoanData$PERIOD_END_LSTAT)
LoanDataCluster <- data.frame(LoanDataCluster,LoanData$grade)

table(LoanDataCluster$LoanData.PERIOD_END_LSTAT,km.out$cluster)
table(LoanDataCluster$LoanData.grade,km.out$cluster)
aggregate(LoanDataCluster[, 1], list(LoanDataCluster$km.out.cluster), mean)
aggregate(LoanDataCluster[, 2], list(LoanDataCluster$km.out.cluster), mean)
aggregate(LoanDataCluster[, 6], list(LoanDataCluster$km.out.cluster), mean)
km.out$size
```

